%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = report.tex

\chapter{Conclusions}
\label{ch:Conclusion}

The original goals of this project, as set forth in the proposal were to:

\begin{description}
    \item[Describe a Failure Model] --- I made some progress in this area, as discussed in \S \ref{section:model:failure}.  Thus, this goal has been accomplished at some level.
    \item[Durability] --- much of the micro-benchmark work that I did was exploring the tension between performance and crash durability.  Thus, I would note that this goal has been accomplished at some level.
    \item[Performance Evaluation] --- this goal was to evaluate adding non-volatile memory support to those systems.  No work was done in this space due to resource constraints.  This remains future work.
\end{description}

The \textbf{primary objective} of this project was to demonstrate the ability to conduct 
independent research.  While I have worked with a number of people during this project, including 
discussing preliminary results and possible future directions, the work described in this document 
rests upon the research that I have conducted.  The experimental results in Chapter \ref{ch:Results}
are directly based upon data that I collected.  I wrote the micro-benchmark code, modified the memory allocator benchmark framework, wrote additional scripts to automate running and collecting a vast volume of data.

Thus, I submit to those evaluating my work that this is sufficient to demonstrate the ability to conduct independent research in pursuit of my PhD at The University of British Columbia.
 